---
title: "The surveydown {{< var rnolink >}} Package"
subtitle: "A flexible, open-source platform for making surveys"
author:
  - "John Paul Helveston, Ph.D."
  - "Pingfan Hu"
date: "Jun 01, 2024"
date-modified: "Jan 28, 2025"
categories: [Software]
execute:
  echo: true
  eval: false
  warning: false
---

## About

{surveydown} is a flexible, open-source survey tool using {{< var rlink >}}, {{< var quarto >}}, {{< var shiny >}}, and {{< var postgres >}}.

<br>
<center>
<a href="https://surveydown.org" class="btn btn-primary" role="button"> Visit surveydown Official Website </a>
</center>

## Awards 🎉

{surveydown} has been awarded the 2nd Prize in the [2024 Student Open Source Awards Program](https://ospo.gwu.edu/2024-student-open-source-award-winners) at GWU!

## Overview

This project is led by [Prof John Helveston](https://www.jhelvy.com). For more details, see our [About](https://surveydown.org/about) page.

<br>

![](/figs/surveydown.svg){fig-align="center" width="550"}

The basic concept is:

1. Design your survey as a {{< var quarto >}} document using markdown and R code.
2. Render your survey into a {{< var shiny >}} app that can be hosted online and sent to respondents.
3. Store survey response data in a {{< var postgres >}} database. We recommend {{< var supabase >}} as a free, secure, and easy to use option.

The {surveydown} R package provides functions to bring this all
together.

**See [Documentation](https://surveydown.org/documentation) to get started making
your own surveydown survey!**

**See [Demos](https://surveydown.org/demos) and use them as your survey templates!**

## Motivation

Most survey platforms (e.g., Google forms, Qualtrics, etc.) use graphic interfaces or spreadsheets to define survey content, making version control, collaboration, and reproducibility difficult or impossible. The surveydown package was designed to address these problems. As an open-source, markdown-based platform, all survey content is defined using **plain text** (markdown and R code) in two files:

- `survey.qmd`: A {{< var quarto >}} document that contains the survey content (pages, questions, etc).
- `app.R`: An R script defining a shiny app that contains global settings (libraries, database configuration, etc.) and server configuration options (e.g., conditional skipping / display, etc.).

This approach makes your survey easy to reproduce, share, and version control with common tools like Git. And since all survey data is stored in a {{< var postgres >}} database, you have total control over where your survey data lives. We provide direct support for {{< var supabase >}} as a free, secure, and easy to use option.

In case you're interested in the background behind the project, this [blog post](https://www.jhelvy.com/blog/2023-04-06-markdown-surveys/) provides something of an origin story. Note that the design discussed in the post is now quite outdated with what ultimately became surveydown.

## Installation

### 1. Install R & Quarto

You need both: 

- Install {{< var rlink >}}
- Install {{< var quarto >}}

We also recommend working with an IDE that has good support for R, Quarto, and Shiny.

{{< var rstudio_ide >}} is great, and we also like {{< var vscode >}} and {{< var positron >}}.

### 2. Install the {surveydown} R package

You can install {surveydown} from CRAN in your R console:

```{r}
#| eval: false

install.packages("surveydown")
```

or you can install the development version from [GitHub](https://github.com/surveydown-dev/surveydown):

```{r}
#| eval: false

# install.packages("pak")
pak::pak('surveydown-dev/surveydown')
```

Load the package with:

```{r}
#| eval: false

library(surveydown)
```

You can also check which version you have installed:

```{r}
#| eval: false

surveydown::sd_version()
```

## License Information

See the license information [here](https://github.com/surveydown-dev/surveydown/blob/master/LICENSE.md).

## Citation Information

If you use this package for in a publication, please cite it! You can
get the citation by typing `citation("surveydown")` into R:

```{r}
citation("surveydown")
```

